TestEvol: A Tool for Analyzing Test-Suite Evolution

### 2012-11-02 icse2013demo-testevol/paper.tex
Revised paper

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. There are several reasons why it is
  important to achieve a thorough understanding of how test cases
  evolve in practice. In particular, researchers who develop automated
  test-repair techniques---an increasingly active research area---can
  use such understanding to develop more effective repair techniques
  that are applicable in real-world scenarios. More generally,
  analyzing test-suite evolution can help testers better understand
  how test cases are modified during maintenance and improve the test
  evolution process, an extremely time consuming activity for any non
  trivial test suite.  Unfortunately, there are no existing tools that
  facilitate investigation of test-suite evolution. To tackle this
  problem, we developed \tool, a tool that enables the systematic
  study of test-suite evolution for Java programs and JUnit test
  cases.  This demo presents \tool and demonstrates its usefulness and
  practical applicability by showing how \tool\ can be used on
  real-world software and unit test suites.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

@@

This demo presents a tool that implements our technique and that we
developed to allow other researchers and practitioners to perform
studies such as we the ones we just summarized. The tool, which is
called \tool, enables a systematic study of test-suite evolution.
Given two versions of a program and corresponding test suites, \tool\
automatically computes differences in the behavior of the test suites
on the two program versions, classifies the actual repairs performed
between the two versions, and computes the coverage attained by the
tests on the versions.

After presenting \tool's features and technical details, the demo
shows how the tool can be used to study how test suites evolved over
the years for a software project.  To do so, we show examples of
application of \tool\ to a set of real-world open-source software
systems.  Specifically, we show how to use \tool\ to investigate
relevant questions on test evolution, such as the following ones: What
types of test-suite changes occur in practice and with what frequency?
How often do test repairs require complex modifications of the tests?
Why are tests deleted and added? Overall, we show how \tool\ is a
useful and practically applicable tool for researchers in the area of
test repair (and test evolution in general), developers and testers
who want to better understand their software maintenance process.

\section{Studying Test-Suite Evolution}
\label{sec:test-evolution}

\subsection{Definitions and Terminology}
\label{sec:terminology}

A \textit{system} $S = (P, T)$ consists of a program $P$ and a test
suite $T$.  A \textit{test suite} $T = \{t_1, t_2, \ldots, t_n\}$
consists of a set of unit test cases.
% A unit test case $t$ is a method (or function) and it is identified
% by its name plus the name of the class it is associated with.
$\testfunc{P}{t}$ is a function that executes test case $t$ on program
$P$ and returns the outcome of the test execution.  A \textit{test
  outcome} can be of one of four types:

@@

\subsection{Test Modifications}
\label{sec:test-mod}

@@

\subsubsection{Category \catrep{} (Repaired Tests)}
\label{sec:category-rep}

@@

\subsubsection{Category \catref{} (Refactored Tests)}
\label{sec:category-c2}

@@

\subsection{Test Deletions}
\label{sec:test-ref}

@@

\subsubsection{Category \catdelaere{} (Hard-To-Fix Tests)}
\label{sec:category-delaere}

@@

\subsubsection{Category \catdelce (Obsolete Tests)}
\label{sec:category-delce}

@@

\subsubsection{Category \catdelp: (Redundant Tests)}
\label{sec:category-delp}

@@

\subsection{Test Additions}
\label{sec:test-add}

@@

\subsubsection{Category \cataddaere: (Bug-Fix Tests)}
\label{sec:category-addaere}

@@

\subsubsection{Category \cataddce (New-Features Tests)}
\label{sec:category-addce}

@@

\subsubsection{Category \cataddp (Coverage-Augmentation Tests)}
\label{sec:category-addp}

@@

\section{Implementation of \tool}
\label{sec:implementation}

\todo{Add the discussion of clone analysis.}

\todo{In general, add more details.}

\todo{Discuss the availability of the tool.}

We implemented our approach for studying test suite evolution in a
tool, called \tool, which facilitates a systematic study of test-suite
evolution for Java programs and JUnit test suites.  \tool\ analyzes a
sequence of versions of a software system, where each version consists
of application code and test code. The versions can be actual releases
or internal builds.  \tool\ consists of five components, as
illustrated in the architecture diagram shown in
Figure~\ref{fig:trex}.

The \emph{compiler} component builds each system version and creates
two jar files, one containing the application classes and the other
containing the test classes.
%
The \emph{test-execution engine} analyzes each pair of system versions
$(S, S')$, where $S = (P, T)$ and $S' = (P', T')$.  First, it executes
$T$ on program $P$ and $T'$ on program $P'$ (\ie it runs the tests on
the respective program versions).  Then, it executes $T'$ on $P$ and
$T$ on $P'$. For each execution, it records the test outcome: $\pass,
\failce, \failae, or \failre$.
%
The \emph{differencing component} compares $T$ and $T'$ to identify
modified, deleted, and added tests. This component is implemented
using the \textsc{wala} analysis infrastructure for Java ({\small
  \url{wala.sourceforge.net}}).

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{architecture.png}
        \vspace*{-16pt}
	\caption{High-level architecture of \tool.}
        \vspace*{-8pt}
	\label{fig:trex}
\end{figure}

@@

Test suites evolve throughout their lifetime and change with the
applications under test. Adapting existing test cases manually,
however, can be extremely tedious, especially for large test suites,
which has motivated the recent development of automated test-repair
techniques (\eg \cite{Daniel:2009, Daniel:2010, Mirzaaghaei:2012}).
To support the development of more effective repair techniques, we
developed \tool, a tool for systematic investigation of test-suite
evolution that can provide its users with a comprehensive
understanding of how test cases evolve. In this demo, we present
\tool, its main features, and its technical characteristics. We also
demonstrate how \tool\ can be run on real-world software systems and
produce a wealth of useful information, such as the types of
test-suite changes that occur, the frequency of the changes, and why
tests are deleted and added.  Overall, we show how \tool\ is a useful
tool not only for researchers working in the area of test repair, and
test evolution in general, but also for developers and testers who
wish to understand their software systems better.

>>>>

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. There are several reasons why it is
  important to achieve a thorough understanding of how test cases
  evolve in practice. In particular, researchers who develop automated
  test-repair techniques---an increasingly active research area---can
  use such understanding to develop more effective repair techniques
  that are applicable in real-world scenarios. More generally,
  analyzing test-suite evolution can help testers better understand
  how test cases are modified during maintenance and improve the test
  evolution process, an extremely time consuming activity for any non
  trivial test suite.  Unfortunately, there are no existing tools that
  facilitate investigation of test-suite evolution. To tackle this
  problem, we developed \tool, a tool that enables the systematic
  study of test-suite evolution for Java programs and JUnit test
  cases.  This demonstration presents \tool and demonstrates its
  usefulness and practical applicability by showing how \tool\ can be
  used on real-world software and unit test suites.
\end{abstract}

@@

\section{Motivation and Overview}
\label{sec:intro}

@@

This demonstration presents a tool that implements our technique and
that we developed to allow other researchers and practitioners to
perform studies such as we the ones we just summarized. The tool,
which is called \tool, enables a systematic study of test-suite
evolution.  Given two versions of a program and corresponding test
suites, \tool\ automatically computes differences in the behavior of
the test suites on the two program versions, classifies the actual
repairs performed between the two versions, and computes the coverage
attained by the tests on the versions.

After presenting \tool's features and technical details, the
demonstration shows how the tool can be used to study how test suites
evolved over the years for a software project.  To do so, we show
examples of application of \tool\ to a set of real-world open-source
software systems.  Specifically, we show how to use \tool\ to
investigate relevant questions on test evolution, such as the
following ones: What types of test-suite changes occur in practice and
with what frequency?  How often do test repairs require complex
modifications of the tests?  Why are tests deleted and added? Overall,
we show how \tool\ is a useful and practically applicable tool for
researchers in the area of test repair (and test evolution in
general), developers and testers who want to better understand their
software maintenance process.

\section{The \tool\ Tool}
\label{sec:test-evolution}

In this section we firs summarize our technique, implemented in the
\tool\ tool, and then discuss the main characteristics of the tool
implementation. A detailed description of the approach can be found
in~\cite{pinto12}.

\subsection{Underlying Technique}

Before describing the characteristics of our technique, we introduce
some necessary terminology.  A \textit{system} $S = (P, T)$ consists
of a program $P$ and a test suite $T$.  A \textit{test suite} $T =
\{t_1, t_2, \ldots, t_n\}$ consists of a set of unit test cases.
% A unit test case $t$ is a method (or function) and it is identified
% by its name plus the name of the class it is associated with.
$\testfunc{P}{t}$ is a function that executes test case $t$ on program
$P$ and returns the outcome of the test execution.  A \textit{test
  outcome} can be of one of four types:

@@

\paragraph*{\textbf{Test Modifications}}
\label{sec:test-mod}

@@

\paragraph{Category \catrep{} (Repaired Tests)}
\label{sec:category-rep}

@@

\paragraph{Category \catref{} (Refactored Tests)}
\label{sec:category-c2}

@@

\paragraph*{\textbf{Test Deletions}}
\label{sec:test-ref}

@@

\paragraph{Category \catdelaere{} (Hard-To-Fix Tests)}
\label{sec:category-delaere}

@@

\paragraph{Category \catdelce (Obsolete Tests)}
\label{sec:category-delce}

@@

\paragraph{Category \catdelp: (Redundant Tests)}
\label{sec:category-delp}

@@

\paragraph*{\textbf{Test Additions}}
\label{sec:test-add}

@@

\paragraph{Category \cataddaere: (Bug-Fix Tests)}
\label{sec:category-addaere}

@@

\paragraph{Category \cataddce (New-Features Tests)}
\label{sec:category-addce}

@@

\paragraph{Category \cataddp (Coverage-Augmentation Tests)}
\label{sec:category-addp}

@@

\section{Implementation}
\label{sec:implementation}

% \todo{Add the discussion of clone analysis.}
% \todo{In general, add more details.}
% \todo{Discuss the availability of the tool.}

In our demonstration, we show an implementation of our technique,
\tool, that works for Java programs and JUnit test suites ({\small
  \url{http://www.junit.org/}}). We chose Java and JUnit because the
former is a widely used language, and the latter is the de-facto
standard unit-testing framework for Java.  \tool\ analyzes a sequence
of versions of a software system, where each version consists of
application code and test code and can be an actual release or an
internal build.

\tool\ is implemented as a Java based web application that runs on
Apache Tomcat ({\small \url{http://tomcat.apache.org/}}). Although the
tool can be installed locally, using a local server, users who want to
get familiar with \tool\ without having to install it can access it
remotely on our servers. The tool is currently available at {\small
  \url{http://cheetah.cc.gt.atl.ga.us:8081/testevol/}}, and
unregistered users can access the tool using account ``icse'', with
password ``icse2013''.

Due to space limitations, we can only provide a few examples of
\tool's graphical interface. The companion video demonstration for
this paper illustrates the tool interface in full detail. Figure...

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{1-summary}
        \vspace*{-16pt}
	\caption{TBD.}
        \vspace*{-8pt}
	\label{fig:summary}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{2-versiondetails}
        \vspace*{-16pt}
	\caption{TBD.}
        \vspace*{-8pt}
	\label{fig:versiondetails}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{3-testdetails}
        \vspace*{-16pt}
	\caption{TBD.}
        \vspace*{-8pt}
	\label{fig:testdetails}
\end{figure}

\tool\ consists of five components, as illustrated in the architecture
diagram shown in Figure~\ref{fig:trex}. 
%
The \emph{compiler} component builds each system version and creates
two jar files, one containing the application classes and the other
containing the test classes.
%
The \emph{test-execution engine} analyzes each pair of system versions
$(S, S')$, where $S = (P, T)$ and $S' = (P', T')$.  First, it executes
$T$ on program $P$ and $T'$ on program $P'$ (\ie it runs the tests on
the respective program versions).  Then, it executes $T'$ on $P$ and
$T$ on $P'$. For each execution, it records the test outcome: $\pass,
\failce, \failae, or \failre$.
%
The \emph{differencing component} compares $T$ and $T'$ to identify
modified, deleted, and added tests. This component is implemented
using the \textsc{wala} analysis infrastructure for Java ({\small
  \url{wala.sourceforge.net}}).

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{architecture}
        \vspace*{-16pt}
	\caption{High-level architecture of \tool.}
        \vspace*{-8pt}
	\label{fig:trex}
\end{figure}

@@

Test suites evolve throughout their lifetime and change with the
applications under test. Adapting existing test cases manually,
however, can be extremely tedious, especially for large test suites,
which has motivated the recent development of automated test-repair
techniques (\eg \cite{Daniel:2009, Daniel:2010, Mirzaaghaei:2012}).
To support the development of more effective repair techniques, we
developed \tool, a tool for systematic investigation of test-suite
evolution that can provide its users with a comprehensive
understanding of how test cases evolve. In this demonstration, we
present \tool, its main features, and its technical
characteristics. We also demonstrate how \tool\ can be run on
real-world software systems and produce a wealth of useful
information, such as the types of test-suite changes that occur, the
frequency of the changes, and why tests are deleted and added.
Overall, we show how \tool\ is a useful tool not only for researchers
working in the area of test repair, and test evolution in general, but
also for developers and testers who wish to understand their software
systems better.
